---
title: "Covid-19 Still a Menace:"
description: |
  An Insight on live data of UK
author:
  - name: Arpan Sarkar
date: 09-02-2021
output:
  distill::distill_article:
    self_contained: false
    toc: True
categories:
  - GGplot
  - Plotly
  - stats
  - highcharts
  - highcharter
  - time-series
  - data-visualization
preview: images/preview.PNG      
---
```{css, echo = FALSE}
.fill {
  color: red;
  font-weight: bold;
}
h2 {
  color: #006DAE;
}

.fill1 {
  color: #c41244;
  font-weight: normal;
}

.fill2 {
  color: #1c2aad;
  font-weight: normal;
}


```



# Historical and live Covid-19 Data Of UK

```{r echo = FALSE, warning = FALSE,message=FALSE}
library(tidyverse)
library(tidycovid19)
library(zoo)
library(plotly)
library(ggforce)
library(patchwork)
library(ggtext)


df <- download_merged_data(cached = TRUE, silent = TRUE)

s<-df %>%
  filter(iso3c == "GBR") %>%
  mutate(
    new_cases = confirmed - lag(confirmed),
    ave_new_cases = rollmean(new_cases, 7, na.pad=TRUE, align="right"),
  new_deaths = deaths-lag(deaths), ave_new_deaths= rollmean(new_deaths, 7, na.pad=TRUE, align="right") ) %>%
  filter(!is.na(new_cases), !is.na(ave_new_cases)) %>%
  filter(!is.na(new_deaths), !is.na(ave_new_deaths))%>%
  ggplot(aes(x = date)) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week",
             date_labels = "%b-%Y")+
          
  labs(y="New cases and deaths")+
  geom_bar(aes(y = new_cases), stat = "identity", fill = "lightblue") +
  geom_line(aes(y = ave_new_cases), color ="orange") +
  geom_line(aes(y= new_deaths), color ="red")+
  theme(axis.text.x = element_text(angle=75))

ggplotly(s)
```


[Covid-19]{.fill} is still threatening the normal life across the globe. 
The new mutant as [[Delta variant]{.fill1}](https://www.cdc.gov/coronavirus/2019-ncov/variants/delta-variant.html) is the new threat.
Efficacy of vaccines are now open beyond doubt.
Watch the current state of affairs with live data from UK. 
From middle of [March 2020]{.fill2}, covid-19 cases started picking up and 
within a month reached a high of [5486]{.fill2} and then tapered down to [below 10]{.fill2} in 2020 July 1st week.  
Again the rise in new cases picked up to more than 33.5K by Nov 12 2020 to fall down again to [1430]{.fill2} by Nov beginning dame year. And menacingly, by the end of 1st week of January 2021, the cases reached the peak with more than [68K]{.fill2}. The cases fell down to [1640]{.fill2} by 10th of April 21.
Again a V shaped  rise in cases to above [47.7K]{.fill2} by July 21, then fell to [25K]{.fill2} in 1st week of August to rise again to [34K]{.fill2} by end of August. 
New fatal cases rose from 1 on 6th March 2020 to [1123]{.fill2} on 10th April 2020. New fatal case fell to 1 by start of August 1st week. Fatal cases rose to above [1450]{.fill2} by start of Feb 2021.On May 3 new death case dropped to 1 again rose very mildly and fell to [0]{.fill2} on 1st June 21. Figure started rising and by end week of August a high of [133]{.fill2} observed.

The startling revelation that is evident from the analysis are:

- Like in other parts of the world post Wuhan outbreak during end of 2019 and start of 2020, UK also got caught by the invasion of the deadly virus Covid-19. March 2020 became the starting of this social evil in the UK with a shocking outcome ending with fatality in many cases. 

- With no solution or medicine available, WHO guidelines lines like Wearing Mask, Sanitizing Hands, Maintain Physical Distance, Avoiding Congregation of More people in public places etc. and then resorting to Lockdowns were practiced. And still being practiced, albeit with lesser degree and periodicity. 
- Fatal cases were high in April 2020 and again in February 2021. The figure then dropped to 0 by 1st June. And the New infected cases also followed similar pattern but with a much higher numbers.

- Prior to January 2021, we find during June to October both fatality and new infections were record low: definitely the effect of strict regulations and lockdowns.

- Since October-November both cases started rising and peak coming in January, reasons being relaxation of lockdowns to give the stalling economy the desired push.

- Big question is How the infection and fatality dropped down from April to July and thereafter new infection increased but fatal cases remain low? Want the answer?  See below.



# Interactive Analysis Of Covid-19 Parameters

```{r echo = FALSE, warning = FALSE,message=FALSE}
library(highcharter)
library(tidyverse)
library(tidycovid19)
library(zoo)
library(plotly)
library(ggforce)
library(patchwork)
library(ggtext)
library(kableExtra)
library(lubridate)
library(summarytools)

df <- download_merged_data(cached = TRUE, silent = TRUE)

gbr<-df %>%
  filter(iso3c == "GBR")

gbr_wide<- gbr %>%  mutate(new_cases = confirmed - lag(confirmed), new_deaths= deaths-lag(deaths),new_vaccination = total_vaccinations - lag(total_vaccinations), new_recovered= recovered- lag(recovered), new_tests= total_tests- lag(total_tests),Month_yr=format_ISO8601(gbr$date, precision = "ym"),vaccination_pcnt=(total_vaccinations/population)*100)%>%select(Month_yr,new_tests,new_cases,new_deaths,new_vaccination,new_recovered,deaths,total_vaccinations,vaccination_pcnt)

gbr_stat<- gbr %>%  mutate(covid_cases = confirmed - lag(confirmed), deaths= deaths-lag(deaths),vaccination = total_vaccinations - lag(total_vaccinations), recovered= recovered- lag(recovered), covid_tests= total_tests- lag(total_tests))%>%select(covid_tests,covid_cases,deaths,vaccination,recovered)

gbr_new<-gbr %>%  mutate(new_cases = confirmed - lag(confirmed), new_deaths= deaths-lag(deaths),new_vaccination = total_vaccinations - lag(total_vaccinations), new_recovered= recovered- lag(recovered), new_tests= total_tests- lag(total_tests))%>%pivot_longer(cols = c("new_cases","new_deaths","new_vaccination","new_recovered","new_tests"), names_to = "variable", values_to = "value")%>% select(date,country, variable,value,hosp_patients,icu_patients)

hchart(gbr_new, "line", 
  hcaes(x = date, y = value, group = variable)
  )

```

Big question is How the infection and fatality dropped down during April to July in 2021 and thereafter new infection increased but fatal cases remain low? Want the answer?  See my next post. 
For the answer, the latest concept of interactive visualisation for multiple parameters on a single canvas is created. Here the users are at liberty to choose parameter(s) of his/her choice. Hovering the cursor over the points of the diagram(s) instantly displays the level and value(s) for ready reference of actual value. Due to scale difference of the relevant parameters, this kind of interactive display is excellent to understand the live behaviour of data.

Coming to the question, yes you have guessed it right. It is the Vaccination that changed the game altogether.

Just to understand it, let's choose two parameters viz. New cases and new vaccination. See the behavior! Second week of January (11/01/2021) the much awaited vaccination started with 1,65,844 persons' inoculation on day one. And during this week the new infection cases were all time high with 68192 on Jan 8, a record high. Gradually, but steadily new infection cases went on dropping below1800 on April 25. It remained below 10k till 20th June. Then onwards started rising to reach another peak with 47848 cases on July 18th. Then dropped a bit, but remained above 25k with 32167 as on August 31st.

Now we need to answer not one question but two questions. Before coming to that let's analyse vaccination data. 

- March 20 showing total inoculation of 844285. Naturally the figure includes 2nd dose cases as well. April 6th vaccination was as low as 272020. Again on May 22nd, total inoculation figure rose to 762361. Again on June 20 vaccination figure dropped to 273158 and thereafter the figures were never below 150000, except on August 30 when it went below 100000 to record low of 86291.

- On January 12th 2021, new death cases were 1246 a d on 13th it was 1568. On 17th figure went down to 682 and on 18th 602 only. It hovered above 1000 but again fell to 594 on 25th. Again little increase but went down to 333 on February 8th. Then little increase but fell to 259 on Feb 14th to 178 on Feb 22nd to 65 only on March 8. Again very little increase, then fell to merely 18 on Mar 22 and remained low and reached to one of the lowest figure of 7 on April 11.

- May 6 figure was only 6, May 10 had only 5, May 16 only 4 and remained below 2 digits to reach the figure of 3 only on June 14. Only after a few days’ figure touched 2 digits but below 20 mostly and again went down to 3 only. Slight increase but June 28 had 3 cases only. After July 12 (figure 6 only) started rising (13th 50 cases, 15th 67 cases, 21st 73 cases, 26th 14 cases, 27th 131 cases, 31st 72 cases, Aug 2nd 25 cases, Aug 10th 146 cases, Aug 17th 170 cases, Aug 24th 174 cases a d Aug 31st 51 cases).

- Now for the prime question, answer is clear. Vaccine worked and is working with desired efficacy. People who were sceptical can now be rest assured on this point.

- The other question that is evident from data, is why from July 2021, cases of infection started rising of a greater proportion although death cases although were little high but not at that proportion? The answer is the advent of new mutant of Covid-19 called Delta Variant for the 1st cause i.e., rise in infection cases and effect of Vaccines for the 2nd cause i.e., fatal cases were contained.

- Statistics of the variables used for analysis are also provided in a tabular format below.


# summary statistics

```{r summary,echo = FALSE, warning = FALSE,message=FALSE}
library(summarytools)

options(scipen = 999)

stat_tab<-descr(gbr_stat)%>%kbl() %>%
  kable_paper("striped", full_width = F) %>%
  kable_material() 

stat_tab

```


# Interactive Data Populated Year-Month Wise

```{r Interactive, echo = FALSE, warning = FALSE,message=FALSE}
library(sparkline)
library(DT)
gbr_wide %>%
  group_by(Month_yr)%>%
  summarise(avg_total_tests = scales::comma(round(mean(new_tests))),
            avg_new_cases = scales::comma(round(mean(new_cases))),
            new_cases_trend = spk_chr(new_cases, type = "line"),
            avg_total_deaths = scales::comma(round(mean(deaths))),
            new_deaths_trend = spk_chr(new_deaths, type = "line")) %>%
 
  
  datatable(escape = FALSE, 
            options = list(columnDefs = list(list(className = 'dt-right', targets = 0:4))),
            colnames = c("Year Month", "Total Tests", "Cases/day", "Covid Cases Trend","Total Deaths", "Deaths/day")
            ) %>%
  spk_add_deps()
```



